---
title: "Class 11: Population Analysis"
author: "Peter Shamasha (A15857589)"
format: pdf
---

## Population Scale Analysis


About ~230 samples have been procesed and and normalizized on a genome level. Now, we want to find whether there is any association between the 4 asthma-associated SNPs on ORMDL3 expression.

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

In order to determine the sample size for genotype, we can use the `table` function and the `$` syntax with the `geno` column in the table in order to table the samples sizes for each genotype.

```{r}
#Sample size of each genotype
table(expr$geno)
```
In order to find the median expression of each of the genotypes, we can make a boxplot using ggplot.


```{r}
library(ggplot2)

ggplot(expr)+aes(x=geno, y=exp)+
  geom_boxplot()
```


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

I will reuse the boxplot from the previous question
```{r}
ggplot(expr)+aes(x=geno, y=exp)+
  geom_boxplot()
```
From this boxplot, we can infer that having the "A" alleles results in higher expression of ORMDL3.
