---
title: "Class 10: Structural Bioinformatics Part 2"
author: "Peter Shamasha (A15857589)"
format: pdf
---

## Structural Bioinformatics Part 2

# Comparative analysis of ADK

ADK (Adenelate kinase) is an important enzyme that maintains the equilibrium between cytoplasmic nucleotides essential for many cellular processes.

We will perform principal component analysis (PCA) on the complete collection of Adenylate kinase structures in the protein data-bank (PDB).


> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

MSA

> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

Grantlab/bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

TRUE


```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```
> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 

214


```{r}
# Blast
#b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
#hits <- plot(b)
```

```{r}
#List out some "top hits"
#head(hits$pdb.id)
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```



Now we can download all these pdf structures
```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Now I want to align and superimpose these structures which are all over the place.

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's Have a look at  our `pdbs` object:

```{r}
pdbs
```

Now we have our aligned and superimposed structutes we can perform all sorts of analysis on them. Let's do PCA...

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```
Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure.


We can cluster the structures by RMSD (or any other method).
```{r}
rd <- rmsd(pdbs)

hc.rd <- hclust(dist(rd))
grps <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps)
```

We can make a wee movie - also called a trajectory of the major differences (i.e. structural displacements) of ADK.

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

